---
title: "Codes"
output:
  html_document:
    toc: yes
    theme: darkly
    toc_float: yes
    number_section: yes
    word_document: defult
  word_document:
    toc: yes
date: "2022-12-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Only  i will write the codes, she explaned about how to connect Rstudio with Git.

How to call library:
```{r echo=TRUE}
library(ggplot2)
library(tinytex)

"how to call function in Library"

"to get working directoy:"
getwd()
"اقدر اكتب بالعربي"
"في نفس الشنك فيه يمين فوق علامه اعدادات فيها خيارات اتحكم في الشنك"
"عدلتها من اعدادت الشنك اني اخليها ينفذ الكود و يظهر لي المخرجات"

```

```{r}
"if i writr like this: ##some thing, the text will be bigger in Knit"

"numberical: 12.5
integer : 5
string: حروف
factor: male
logical: True,false"


```


```{r}
x <- 5
typeof(x)
y <- "Omar"
typeof(y)

A <- "55"
typeof(A)

is.character(y)
is.double(y)
Q <- FALSE
typeof(Q)
w <- 2/0

w


```
# use c , type of, 

```{r}
v <- c(1,2.5,4,6)
typeof(v)

M <- c(2.2,3,"A","b")
typeof(M)
"الأولويه للأسترنج"

DS <- c(TRUE,3,7.2)
typeof(DS)

```
# sequence data and using ?,?? fro exploaring libraryes
```{r}
"عشان اعرف عن مكتبه و طريقه استخدامها لازم اكتب استفهام و اسم المكتبه  في الكونسول الي تحت هذي النافذه"
?seq
??seq


DSD <- seq(1:20)
FDS <- seq(4,20, by=2)


```
#change the data type,"the below code make me able to change the data type from double to integer, just use: as.type of data i want"
```{r}

typeof(5)

as.integer(x)

Dateee <- 1-1-2023
typeof(Dateee)
as.character.Date(Dateee)
```

#explain
```{r}

"
3**9 :
ثلاثه اوس تسعه
Sqrt(9) = 3 :
الجذر
3+3*5 = 18
هنا الاولوليه للضرب
اذا حاب اجمع او ف اسوي كذى:
(3+5)*3
-
Titoruial ((تدريب عملي من نفس الار ستوديو ، يكون محله جمب المتغيرا الي نعرفه فوق يمين

"

```

#explain
```{r}
"here pacage: it is make me able to open my project from multible devices with out mistakes in working directory
"


```

##import data
```{r}
library(tidyverse)
library(here)
```

```{r}
read_csv(here("data","SAFI_clean.csv"))
Datafile <- read.csv(here("data","SAFI_clean.csv"), na = "NULL")

```
#"data type:"

```{r}
glimpse(Datafile)

Datafile$no_membrs

no_of_member <- Datafile$no_membrs

view(no_of_member)
"showed what is unique in the data"
unique(no_of_member)
unique(Datafile)

"length: show as number how many unique value:"
length(unique(no_of_member)) 
length(unique(Datafile))

Datafile %>% 
  select(village,no_membrs, months_lack_food)

Only <- Datafile %>% 
  select(village,no_membrs, months_lack_food)
"this code below with -c means delet the selected data"
delet_few <- Datafile %>% 
  select(-c(village,no_membrs))

"this code below with -c means delet the selected data, using : to select From .. To .."

delet_few_fwe <- Datafile %>% 
  select(-c(village:no_membrs))
"using select with filter to chose what i need from the dataset"
SSelect <- Datafile %>% 
  select(village:years_liv) %>% 
  filter(village == "God",
         years_liv > 24)

Datafile$village=="God"

```

```{r}
samlldata <- Datafile %>% 
  select(village:respondent_wall_type,rooms) %>% 
  filter(village == "God" & rooms > 1 )


Task <- Datafile %>% 
  select(memb_assoc,affect_conflicts,liv_count,no_meals) %>% 
  filter(memb_assoc == "yes")

filter(Datafile, memb_assoc == "yes")
```
```{r}
"mutate to generat a new column with our needs from the data"

Datafile %>% 
  mutate(pepole_per_room = no_membrs / rooms)

"
it is function that change the number to integer:
round()
"
newd1<- Datafile %>% 
  mutate(pepole_per_room = round(no_membrs / rooms))

```

```{r}
"
this function showed to me the NA using factr yes or no
is.na(data$the_column_name)
TRUE= NA, FALSE = Real value
"

is.na(Datafile$memb_assoc)
"to delete missing vlaue"

misss <- Datafile %>% select(memb_assoc)

clean_miss <- misss %>% select(memb_assoc) %>% 
  filter(!is.na(memb_assoc))
"
if i used is.na with out ! = it will delete the real value and keep missing value"


```

```{r}
"after cleaning"
after_clean_and_filter<- Datafile %>% 
  filter(!is.na(affect_conflicts)) %>% 
  mutate(pepole_per_room = round(no_membrs / rooms))
"i will see the data in after_clean_and_filter less than the original data becuse i delet the missing value and the result for the filter is right."



```


```{r}
after_clean_and_filter <- after_clean_and_filter %>% 
  mutate(Total_meal = no_membrs * no_meals) %>% 
  filter(Total_meal > 20)

after_clean_and_filter

view(after_clean_and_filter)

"another way to write mutate"

grouped <- Datafile %>% 
  group_by(village) %>% 
  summarise(mean = mean(no_membrs), max= max(no_membrs)) %>% 
  ungroup()

view(grouped)
```


```{r}
"code from me: How to creat data in table with spesific values"
dff <- data.frame(
  Name = c ("a", "B", "C"),
  Age = c (5, 42, 37)
)

```
```{r}
"Usualy we use summary after group by"
Datafile %>%
  group_by(no_meals) %>% 
    summarise(count = n())



Datafile %>%
  group_by(no_meals) %>% 
    count()

```

import ggplot
```{r}
library(ggplot2)
```
# ggplot
```{r}
ggplot(Datafile, aes(x=no_of_member, y=no_of_member))+
  geom_point() +
  geom_smooth() +
  coord_flip() +
  theme(legend.position = "none")

Datafile
```

#Data from R studio to train
```{r}
data()
```

```{r}
library(gapminder)
```

```{r}
gapData <- gapminder
```
#import the data from the web
```{r}
ca <- read_csv("https://raw.githubusercontent.com/scienceParkStudyGroup/r-lesson-based-on-ohi-data-training/gh-pages/data/ca.csv")
ca
```
#exploaring data method
```{r}
glimpse(ca)
str(ca)
```
#st plot
```{r}
ggplot(
  
  ca, aes(x=year, y= visitors, colour = park_name, size =region )
) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  theme(legend.position = "none") +
  labs(x="year", y="visitors", title="Californai national park visitation")+
  theme_minimal() +
  theme(
    legend.title = element_blank()
    
  )
```

```{r}
ggplot(
  
  ca, aes(x=year, y= visitors, colour = park_name, size =region )
) +
  geom_smooth() +
  facet_wrap(~ code, scales = "free") + 
  theme_bw()
```

```{r}
"R studio dral with arabic languge with xlsx bettwe than csv"

ca$state
unique(ca$state)
"اضافه اسم الي ملف الى ملف 
ignor
في خانه الملفات يمين تحت يخليه ماينرفع الى الجت هب حتى لو سويت كوممت"
view(gapminder)
```

```{r}
  ggplot(gapminder, aes(x=log(gdpPercap),col= year, y= lifeExp, size =pop )
) +
  geom_point(alpha = 0.5, color = "red" )+
  geom_smooth(method = lm)+
  facet_wrap(~ continent, scale = "free")+
  theme_bw()
```
"install.packages("ggplotAssist")
بعد م اثبتها بتطلع لي في 
Addins
اعلى شي فووق تحت 
Help
اختار منها 
ggplot Assit
 ثم بتفتح لي رابط مثل
 192.186.1.1
 فيها صفحه  تساعدني تماما على 
 انشاء 
 ggplot
 للداتا عن طريق يوزر انتر فيس، و اعرف عن طريقها كل خصائص ال
 ggplot
 
"
```{r}

```

#"plotly
هي مكتبه تول الاشكال البلوتات الى اشكال تفاعليه
ولكن عشان تشتغل 
لازم
ادخل كود البلوت في
<- 
ثم ادخله في كود
ggplotly(اسم الي  حطيته)

", "How to save plot and the output i made, dpi = وضوح وجوده الصوره, hight = "auto", width = "auto""

"ggplotly(pplloott)
"
```{r}
 pplloott<- ggplot(gapminder, aes(x=log(gdpPercap),col= year, y= lifeExp, size =pop )
) +
  geom_point(alpha = 0.5, color = "red" )+
  geom_jitter()+ 
  geom_smooth(method = lm)+
  theme_bw()


ggsave(here("results","h0ere.png"), pplloott, dpi = 300 )
library(plotly)
```

```{r}
"
rstat
اذا ابغا ادور عن اي شي في الار اكتب اسم الشي الي ابغا ادور عليه ثم اكتب هذى الاسم
مثلا:
plotly rstat install
,,
اذا ابغا اعرف عن اي شي 
اكتب اسمه في قوقل و اكتب
sheetcheat.
"

"https://rweekly.org/
this is source from learning and being collabiraive with R community"
```



#How to save data after minpulation
```{r}
write_csv(ca, here("results","ca.csv"))
```

```{r}
"changes in Yaml: which is in the top of this page"
```
#"https://www.datadreaming.org/post/r-markdown-theme-gallery/" "this links has muti thems i can use in YAML ( in the top of this page)", to publish my paper in internet for free,  1- kint the file 2- upside right ,3- RPubs then publsh.
```{r}
```
# library(prettydoc): it is used for design the yaml  and it has a lot of thems i can check the website: "https://www.datadreaming.org/post/r-markdown-theme-gallery/" to know more about it

```{r}

```
#ymlthis: it used to write yml code so need it to wrok with out errors becuse yml is sensetive to the spaces, after activate ymlthis lib it will show in Addins in the upper side of the Rstudio.

```{r}
library(ymlthis)

```

#"the other way to creat them is : file - new filre - R mark down - then i will choose them - select any theme i like"
```{r}

```


