---
title: "Distill codes"
description: |
  A short description of the post.
author:
  - name: Nora Jones
    url: https://example.com/norajones
date: 2023-01-03
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.

#for creating Distill post: in the console down below, write distill::creat_post("title the post"), then the content here. "If i want to write in about.html i need to write in it and if i want to upload images i need to put the image in the same folder for exampile: i want to upload image in about me, the app will creat folder to me names as *about.html* then i need to put the image in this folder becuse i want to upload image in it."
```{r}
library(tidyverse)
```

#2 to upload this file to Github i need to use library (usethis), then insert "usethis::use_git()" then the package will as :Is it ok to commit it?
i chose 1 as yes, Done.
```{r}

```
#with each changes in this pages we need to use Knit or From Build upper side right click Build Website.
```{r}

```

