---
title: "Session 9"
author: "Batool Almarzouq"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework

The answer as Anhal completed:

```{r}
## libs
library(tidyverse)
library(tidytuesdayR)
library(gapminder)
library(here)
```

```{r}
## read the dataset
coffee_list <- tidytuesdayR::tt_load('2020-07-07')

```

```{r}
## convert from list to datframes
coffee <- do.call(rbind,coffee_list)
coffee_list$coffee_ratings
```

```{r}

coffee %>%
  count(variety)
```


```{r}

co2 <- coffee %>%
  mutate(variety = fct_lump(variety, 10))
```


If you are new to `fct_lump()`, you can go through [this lesson](https://cran.r-project.org/web/packages/forcats/vignettes/forcats.html)

```{r}
## simple manipluation of the data
co <- coffee %>%
  filter(!is.na(variety), total_cup_points > 10) %>%
  mutate(variety = fct_lump(variety, 10),sort = TRUE)

```

```{r}
## plot the data 
co %>%
 ggplot( aes(total_cup_points, variety ))+
  geom_boxplot()
```

## Save plots and datasets

```{r}

plot <- gapminder %>% 
  ggplot(aes(x = log(gdpPercap), y = lifeExp, col = year, size = pop)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = lm) +
  facet_wrap( ~ continent, scales = "free") +
  theme_bw()

gapminder_data <- gapminder

```

```{r}
#ggsave(here("results","plot.png"), plot, width = 15, height = 10)

#write_csv(gapminder_data, here("results","gapminder.csv"))
```

# Maniplutaing YAML in Markdown

1- toc

2- floating toc

3- [themes](https://www.datadreaming.org/post/r-markdown-theme-gallery/)


4- Changing the output


## Use yamlthis to create the header

ymlthis is an R package intended to make it easier for you to generate YAML headers

`ymlthis:::launch_yaml_addin()`

## Publish with RPub

Easy web publishing from R. Write R Markdown documents in RStudio. Share them here on RPubs.

## Exc

Install prettydoc package and change the YAML header to:

    output:
     prettydoc::html_pretty:
       theme: cayman
       highlight: github
       math: katex

Publish your page in RPub?

# Using renv

Underlying the philosophy of renv is that any of your existing workflows should just work as they did before -- renv helps manage library paths (and other project-specific state) to help isolate your project's R dependencies.

The general workflow when working with renv is:

-   Call `renv::init()` to initialize a new project-local environment with a private R library,

Work in the project as normal, installing and removing new R packages as they are needed in the project,

-   Call `renv::snapshot()` to save the state of the project library to the lockfile (called renv.lock),

Continue working on your project, installing and updating R packages as needed.

-   Call `renv::snapshot()` again to save the state of your project library if your attempts to update R packages were successful, or call renv::restore() to revert to the previous state as encoded in the lockfile if your attempts to update packages introduced some new problems.

-   `renv::restore()` to restore the state of your project from renv.lock.

## Joining Datasets

```{r}
#create dataframes

demographics <- data.frame(
  id = c("P1", "P2", "P3"),
  age = c(40, 54, 47),
  country = c("GER", "JPN", "BRA")
  )

adverse_events <- data.frame(
  id = c("P1", "P1", "P3", "P4"),
  term = c("Headache", "Neutropenia", "Constipation", "Tachycardia"),
  onset_date = c("2020-12-03", "2021-01-03", "2020-11-29", "2021-01-27")
  )
```

```{r}
left_join <- left_join(demographics, adverse_events, by = "id")

```

```{r}
inner_join <- inner_join(demographics, adverse_events, by = "id")

```

```{r}

right_join(demographics, adverse_events, by = "id")

full_join(demographics, adverse_events, by = "id")

```

## Creating issues in GitHub

```{r}
sessionInfo()
```
