---
title: "Session6_Fatimah"
author: "Fatimah"
date: '2022-07-14'
output: html_document
---

## Import Data

```{r}
library(tidyverse)
library(here)

```

```{r}

interviews <- read_csv(here("data","SAFI_clean.csv"), na = "NULL")

```


```{r}
glimpse(interviews)

wall_type <- interviews$respondent_wall_type

length(unique(wall_type))

```


```{r}

small_data <- interviews %>%
  select(village:respondent_wall_type, rooms) %>% 
  filter(village == "God" & rooms > 1)
  

```
# Ex1
Using pipes, subset the interviews data to include interviews where respondents were members of an irrigation association (memb_assoc) and retain only the columns affect_conflicts, liv_count, and no_meals
```{r}
small_data3 <- interviews %>% 
  select(memb_assoc,affect_conflicts,liv_count,no_meals) %>%
  filter(memb_assoc == "yes")
```

### Add new column
##### **round** to remove the dicemels

```{r}
small_data4 <- interviews %>% 
  mutate(people_per_room = round(no_membrs / rooms))
```

```{r}
is.na(interviews$affect_conflicts)
```


### to filter NA

```{r}
small_data4.1 <- interviews %>% 
  filter(!is.na(affect_conflicts)) %>% 
  mutate(people_per_room = round(no_membrs / rooms))
  
```
# Ex2
Create a new dataframe from the interviews data that meets the following criteria:
  contains a new column called total_meals containing a value that is equal to the total number of meals served in the household per day on average (no_membrs times no_meals). 
Include only the rows where total_meals is greater than 20 should be shown in the final dataframe.
```{r}
small_data5 <- interviews %>% 
  mutate(total_meals = no_meals * no_membrs) %>% 
  filter(total_meals > 20)
```

## Group by specific character >>

```{r}
small_data6 <-  interviews %>% 
  group_by(village) %>% 
  summarise(mean(no_membrs))

```

```{r}
small_data6 <-  interviews %>% 
  group_by(village) %>% 
  summarise(mean=mean(no_membrs), max= max(no_membrs)) %>% 
  ungroup()

```

